---
title: "GDP"
author: "Daning Bi"
date: "Febuary 21, 2016"
output: pdf_document
---
# Part a

First of all we need to read the data into R:

```{r}
setwd("D:\\ANU\\2016S1\\STAT8027\\stat8027\\")
gdp <- read.table("GDP.txt", header = T)
```

Then before we plot the data, we should take natural log of GDP fisrt as the scale of data is quite big which may affect the plot:

```{r}
gdp$Y2013<-log(gdp$Y2013)
```

After transformation, we can take a plot as well as a histogram of the log gdp:
```{r}
par(mfrow=c(1,2))
plot(gdp$Country.Code,gdp$Y2013, main = "Plot of logGDP")
hist(gdp$Y2013, main = "Histogram of logGDP")
```

From the picture we may found that the log GDP of different countries are close to a normal distribution and the plot shows that the log GDPs are almost randomly allocated around 25.

# Part b

Next we can compute a six number summary of log GDP which includes mean, median, maximum, minimum, first quantile and third quantile.

```{r}
summary(gdp$Y2013)
```

And we can also found that there are 39 missing values in log GDP.

# Part c

In addition, we can also take a look of box plot to check if there are any outliers or whether the log GDP is skewed:

```{r}
boxplot(gdp$Y2013)
```

From the boxplot we can conclude that there is a outlier lies outside the range (1st quartile - 1.5 IQR, 3rd quartile + 1.5 IQR), which is the smallest number.
And the data is slightly skewed to the left (i.e. negatively skewed)
Therefore, there is one value smaller than 1st quartile - 1.5 IQR and all the value are smaller than 3rd quartile + 1.5 IQR. We can locate the outlier using the following code:

```{r}
gdp[which.min(gdp$Y2013),]
```

Hence we found that Tuvalu had the smallest GDP in 2013.

# Part d

The best guess for $\mu$ and $\sigma$ are $T_1 = \frac{1}{N}\sum\limits_{i=1}^{N} Y_i$ and $T_2 = \frac{1}{N-1}\sum\limits_{i=1}^{N} (Y_i-\bar{Y})^2$.
The expected values of $T_1$ and $T_2$ are just $\mu$ and $\sigma^2$, as both of them are unbiased estimates. $T_1$ and $T_2$ and be found using the following codes.
```{r}
mean(gdp$Y2013, na.rm = T)
var(gdp$Y2013, na.rm = T)
```