---
title: "GDP2"
author: "Daning Bi"
date: "March 3, 2016"
output: pdf_document
header-includes:
- \usepackage{bbm}
- \usepackage{amsmath}
---
# Part a

First of all we need to read the data into R:

```{r}
setwd("D:\\ANU\\2016S1\\STAT8027\\stat8027")
gdp <- read.csv("gdp2013.csv")
labor <- read.csv("labor2013.csv")
```

Before we run the linear regression model, we need to have a check of the data. We take logritham to remove the skewness of the data, then plot the labor with gdp. The missing value has also been removed.

```{r}
y<-as.matrix(na.omit(log(gdp[,3])))
g<-as.matrix(na.omit(log(labor[,3])))

plot(g,y)

```

According to the matrix form solution of linear regression, $
$$\boldsymbol{\beta =(X'X)^{-1}X'Y}$$.

Hence we can solve the linear regression parameters using matrix calculation:
```{r}
i <- as.matrix(rep(1,length(g)))
x <- cbind(i,g)
beta <- solve(t(x)%*%x)%*%t(x)%*%y
beta
```

Hence we can solve that $\boldsymbol\beta_{0}=19.447$, and $\boldsymbol\beta_{1}=0.353$.

Then we can use `lm()` function in R to check if the answer we calculated is correct or not:

```{r}
lr <- lm(y~g)
summary(lr)
lr$coefficients
```

As they are exactly the same, we have double checked our solution.